Time invariant analysis of epidemics with \pkg{EpiCompare}

### 2021-05-22 jss/EpiCompare-paper/EpiCompare-paper.Rmd
step 1 of edits to section 3

\shannon{paragraph re-write thoughts: I changed this paragraph to align with the other sub-sections.  We first talk about the step in the pipeline and then how epicompare fits into that pipeline.  The old paragraph skipped the general plots, tables, and 'standard' tools and went straight into filaments.  I re-wrote the paragraph to more clearly indicate that we don't think filaments are the only tools for comparison and assessment.}

Finally, the data analysis pipeline often ends with plots, tables, and summary statistics that are used to compare \shannon{and assess} different models, and simulations to one another\footnote{\ben{[Ben says: I'm confused why this is just about comparison and not assessment. Also, I didn't really want to make tons of comments on this section, so could you add a footnote on why you rewrote it and what you were going for?]}}. In \pkg{EpiCompare} we provide a set of comparison and assessment tools for model and simulation results that extend beyond the standard performance metrics (e.g. mean squared error or AIC) and into the lens of time-invariant analysis.\footnote{\shannon{a bit of an awk transition but I wanted to use past wording about TIA.}}  We have found these tools to be specifically applicable for situations where only one season or cycle of an epidemic has occurred.

@@

@@

>>>>

\ben{\sout{Finally,} Often}\footnote{\ben{[Ben says: I feel like starting with "finally" causes a weird pause - but maybe that's just me.]}} the data analysis pipeline \ben{\sout{often}} ends with plots, tables, and summary statistics \ben{\sout{that} which} are used to \sout{\ben{compare} \shannon{and assess}} \ben{\sout{different models, and simulations to one another} assess model preformance and compare across models or simulations.} In \pkg{EpiCompare} we provide a set of comparison and assessment tools for model and simulation results that extend beyond the standard performance metrics (e.g. mean squared error or AIC) and into the lens of time-invariant analysis.  We have found \ben{that} these tools \ben{\sout{to be} are} specifically applicable for situations where only one season or cycle of an epidemic has occurred, \ben{or is the object of interest}.

\ben{The first set of of tools surround the creation of prediction regions. We can create a prediction regions from model simulations to examine if our model simulations capture the true epidemic trajectory. We do so in a time-invariant way and utlize a filamental representation of the model simulations and the true epidemic.} [For three-state epidemic models, we provide the `ggplot`/`ggtern` extension `geom_prediction_band()` which creates a prediction region around the top $1-\alpha$ proportion of the simulations. In this visual setting, comparing this prediction to the true epidemic trajectory or comparing the prediction regions defined by two different models' simulations can be done by eye. In \pkg{EpiCompare}, we also provide these prediction regions for epidemic models with with more than three states. The functions `create_convex_hull_structure()` and `create_delta_ball_structure()` create different geometric representations of prediction regions for any dimensional state-based model. For both of these geometric structures, we provide functions to check if a path is contained (`contained()`)]. \ben{We can also use these prediction regions to visually or mathematically compare how similar two sets of simulations are. In \pkg{EpiCompare} we provide the} `hausdorff_dist()` \ben{function to calculate the Hausdorff distance between multiple prediction regions.}

[We also provide functions to calculate the "extremeness" of a true epidemic trajectory compared to simulated epidemics via the equi-distance filamental trajectory representation as mentioned in Section \ref{sec:beyond-r0-sir}.] \ben{We provide implementations of a few score functions that capture extremeness of epidemics relative to other epidemics, and these scores can be turned into an extremeness measure with} `mean(sim_scores > truth_score)`. [Specifically, functions like  `distance_pseudo_density_function()` can calculate a pseudo-density estimate of the true epidemic relative to simulated ones. Functions `distance_depth_function()` and `local_distance_depth_function()` provide depth scores that suggest how geometrically central an epidemic is to simulations.]

]]

[[\textcolor{red}{ OLD TEXT}

@@

We also provide functions to calculate the "extremeness" of a true epidemic trajectory compared to simulated epidemics via the equi-distance filamental trajectory representation as mentioned in Section \ref{sec:beyond-r0-sir}. Specifically, functions like  `distance_pseudo_density_function()` can calculate a pseudo-density estimate of the true epidemic relative to simulated ones. Functions `distance_depth_function()` and `local_distance_depth_function()` provide depth scores that suggest how geometrically central an epidemic is to simulations.

@@

[[\textcolor{red}{Really OLD TEXT}

### 2021-05-22 jss/EpiCompare-paper/EpiCompare-paper.tex
step 1 of edits to section 3

\textcolor{orange}{paragraph re-write thoughts: I changed this paragraph to align with the other sub-sections.  We first talk about the step in the pipeline and then how epicompare fits into that pipeline.  The old paragraph skipped the general plots, tables, and 'standard' tools and went straight into filaments.  I re-wrote the paragraph to more clearly indicate that we don't think filaments are the only tools for comparison and assessment.}

Finally, the data analysis pipeline often ends with plots, tables, and
summary statistics that are used to compare
\textcolor{orange}{and assess} different models, and simulations to one
another\footnote{\textcolor{violet}{[Ben says: I'm confused why this is just about comparison and not assessment. Also, I didn't really want to make tons of comments on this section, so could you add a footnote on why you rewrote it and what you were going for?]}}.
In \pkg{EpiCompare} we provide a set of comparison and assessment tools
for model and simulation results that extend beyond the standard
performance metrics (e.g.~mean squared error or AIC) and into the lens
of time-invariant
analysis.\footnote{\textcolor{orange}{a bit of an awk transition but I wanted to use past wording about TIA.}}
We have found these tools to be specifically applicable for situations
where only one season or cycle of an epidemic has occurred.

@@

@@

>>>>

\textcolor{violet}{\sout{Finally,} Often}\footnote{\textcolor{violet}{[Ben says: I feel like starting with "finally" causes a weird pause - but maybe that's just me.]}}
the data analysis pipeline \textcolor{violet}{\sout{often}} ends with
plots, tables, and summary statistics
\textcolor{violet}{\sout{that} which} are used to
\sout{\textcolor{violet}{compare} \textcolor{orange}{and assess}}
\textcolor{violet}{\sout{different models, and simulations to one another} assess model preformance and compare across models or simulations.}
In \pkg{EpiCompare} we provide a set of comparison and assessment tools
for model and simulation results that extend beyond the standard
performance metrics (e.g.~mean squared error or AIC) and into the lens
of time-invariant analysis. We have found \textcolor{violet}{that} these
tools \textcolor{violet}{\sout{to be} are} specifically applicable for
situations where only one season or cycle of an epidemic has occurred,
\textcolor{violet}{or is the object of interest}.

\textcolor{violet}{The first set of of tools surround the creation of prediction regions. We can create a prediction regions from model simulations to examine if our model simulations capture the true epidemic trajectory. We do so in a time-invariant way and utlize a filamental representation of the model simulations and the true epidemic.}
{[}For three-state epidemic models, we provide the
\texttt{ggplot}/\texttt{ggtern} extension
\texttt{geom\_prediction\_band()} which creates a prediction region
around the top \(1-\alpha\) proportion of the simulations. In this
visual setting, comparing this prediction to the true epidemic
trajectory or comparing the prediction regions defined by two different
models' simulations can be done by eye. In \pkg{EpiCompare}, we also
provide these prediction regions for epidemic models with with more than
three states. The functions \texttt{create\_convex\_hull\_structure()}
and \texttt{create\_delta\_ball\_structure()} create different geometric
representations of prediction regions for any dimensional state-based
model. For both of these geometric structures, we provide functions to
check if a path is contained (\texttt{contained()}){]}.
\textcolor{violet}{We can also use these prediction regions to visually or mathematically compare how similar two sets of simulations are. In \pkg{EpiCompare} we provide the}
\texttt{hausdorff\_dist()}
\textcolor{violet}{function to calculate the Hausdorff distance between multiple prediction regions.}

{[}We also provide functions to calculate the ``extremeness'' of a true
epidemic trajectory compared to simulated epidemics via the
equi-distance filamental trajectory representation as mentioned in
Section \ref{sec:beyond-r0-sir}.{]}
\textcolor{violet}{We provide implementations of a few score functions that capture extremeness of epidemics relative to other epidemics, and these scores can be turned into an extremeness measure with}
\texttt{mean(sim\_scores\ \textgreater{}\ truth\_score)}.
{[}Specifically, functions like
\texttt{distance\_pseudo\_density\_function()} can calculate a
pseudo-density estimate of the true epidemic relative to simulated ones.
Functions \texttt{distance\_depth\_function()} and
\texttt{local\_distance\_depth\_function()} provide depth scores that
suggest how geometrically central an epidemic is to simulations.{]}

{]}{]}

{[}{[}\textcolor{red}{ OLD TEXT}

@@

We also provide functions to calculate the ``extremeness'' of a true
epidemic trajectory compared to simulated epidemics via the
equi-distance filamental trajectory representation as mentioned in
Section \ref{sec:beyond-r0-sir}. Specifically, functions like
\texttt{distance\_pseudo\_density\_function()} can calculate a
pseudo-density estimate of the true epidemic relative to simulated ones.
Functions \texttt{distance\_depth\_function()} and
\texttt{local\_distance\_depth\_function()} provide depth scores that
suggest how geometrically central an epidemic is to simulations.

@@

{[}{[}\textcolor{red}{Really OLD TEXT}
